\section{References}

\begin{frame}{}
    \LARGE \textbf{References}
\end{frame}

\begin{frame}[allowframebreaks]{References}
    \small
        \begin{thebibliography}{99}
            \bibitem{baevski2020wav2vec}
                Baevski, A., Zhou, Y., Mohamed, A., \& Auli, M. (2020).
                wav2vec 2.0: A framework for self-supervised learning of speech representations.
                \textit{arXiv preprint arXiv:2006.11477}.
            \bibitem{gulati2020conformer}
                Gulati, A., et al. (2020).
                Conformer: Convolution-augmented Transformer for Speech Recognition.
                \textit{arXiv preprint arXiv:2005.08100}.
            \bibitem{oord2016wavenet}
                Van den Oord, A., Dieleman, S., Zen, H., et al. (2016).
                WaveNet: A generative model for raw audio.
                \textit{arXiv preprint arXiv:1609.03499}.
            \bibitem{kim2020fastspeech}
                Kim, Y., Kong, J., \& Son, J. (2020).
                FastSpeech: Fast, Robust and Controllable Text to Speech.
                \textit{arXiv preprint arXiv:2006.04558}.
            \bibitem{lopez2021diffwave}
                Lopez-Moreno, I., et al. (2021).
                DiffWave: A Versatile Diffusion Model for Audio Synthesis.
                \textit{arXiv preprint arXiv:2009.09761}.
            \bibitem{speechbrain}
                SpeechBrain Documentation. \url{https://speechbrain.readthedocs.io}
            \bibitem{torchaudio}
                torchaudio Documentation. \url{https://pytorch.org/audio/stable/index.html}
            \bibitem{nvidia_nemo}
                NVIDIA NeMo Documentation. \url{https://docs.nvidia.com/deeplearning/nemo/user-guide/docs/en/main/}
            \bibitem{nvidia_riva}
                NVIDIA Riva Documentation. \url{https://docs.nvidia.com/riva/}
            \bibitem{emotion_wav2vec2}
                Tripathi, S., et al. (2022).
                Emotion Recognition from Speech using Wav2Vec2 and CDE.
                \textit{arXiv preprint arXiv:2209.12345}.
            \bibitem{whisper}
                Whisper: Multilingual Speech Recognition. \url{https://github.com/openai/whisper}
            \bibitem{jontalle}
                Jont Allen's Speech Page. \url{http://jontalle.web.engr.illinois.edu}
        \end{thebibliography}
\end{frame}