\begin{frame}[allowframebreaks]{Practical Implementations and Considerations}
    \begin{itemize}
        \item \textbf{Key Architectures and Models}
        \begin{itemize}
            \item \textbf{Classic Approaches:}
            \begin{itemize}
                \item \textbf{Bayesian Networks:} Graphical models that encode conditional dependencies.
                \item \textbf{Recurrent Neural Networks (RNNs):} Naturally suited for sequential data modeling.
            \end{itemize}
            \item \textbf{Modern Neural Approaches:}
            \begin{itemize}
                \item \textbf{Masked Autoencoder for Distribution Estimation (MADE):} A feed-forward neural network with masks to enforce the autoregressive property.
                \item \textbf{Causal Masked Neural Models:} Neural networks designed for autoregressive generation by preventing information flow from future elements.
                \begin{itemize}
                    \item \textbf{Convolutional:} PixelCNN, WaveNet (use masked convolutions).
                    \item \textbf{Attention-based:} \textbf{Transformers} with causal masks (enable flexible context modeling).
                \end{itemize}
            \end{itemize}
        \end{itemize}
    \end{itemize}

    \framebreak

    \begin{itemize}
        \item \textbf{Design Considerations}
        \begin{itemize}
            \item \textbf{Tokenization:} Discretizing continuous data (e.g., images into patches, audio into samples) or using natural discrete units (words, bytes) to form sequences for modeling.
            \item \textbf{Caching:} Essential for efficient inference and generation in autoregressive models, by reusing previously computed states.
            \item \textbf{Architecture Choices:}
            \begin{itemize}
                \item \textbf{Decoder-only vs. Encoder-Decoder:} Trade-offs exist for generation and conditioning capabilities.
                \item \textbf{Newer Recurrent Models:} Evolution of RNNs and hybrid architectures continue to emerge.
            \end{itemize}
        \end{itemize}
    \end{itemize}
\end{frame}