\begin{frame}[allowframebreaks]{References}

\begin{thebibliography}{9}

\bibitem{cs231n}
Fei-Fei Li, Yunzhu Li, and Ruohan Gao. 
\textit{Stanford CS231n: Deep Learning for Computer Vision}. 
\url{http://cs231n.stanford.edu/index.html}

\bibitem{dl4cv}
Assaf Shocher, Shai Bagon, Meirav Galun, and Tali Dekel. 
\textit{WAIC DL4CV Deep Learning for Computer Vision: Fundamentals and Applications}. 
\url{https://dl4cv.github.io/index.html}

\bibitem{umich}
Justin Johnson. 
\textit{UMich EECS 498.008/598.008: Deep Learning for Computer Vision}. 
\url{https://web.eecs.umich.edu/~justincj/teaching/eecs498/WI2022/}

\bibitem{vaswani2017attention}
Ashish Vaswani, Noam Shazeer, Niki Parmar, Jakob Uszkoreit, Llion Jones, Aidan N Gomez, Łukasz Kaiser, and Illia Polosukhin.
\textit{Attention Is All You Need}.
Advances in Neural Information Processing Systems, 2017.
\url{https://arxiv.org/abs/1706.03762}

\bibitem{cs231n-attention}
Stanford CS231n.
\textit{Lecture 11: Attention \& Transformers}.
\url{http://cs231n.stanford.edu/slides/2023/lecture11.pdf}

\bibitem{cmu-dl-transformers}
Carnegie Mellon University.
\textit{Deep Learning Lecture: Vision Transformers and Transformers}.
\url{https://deeplearning.cs.cmu.edu/F23/index.html}
\url{https://www.youtube.com/watch?v=1gq2y1E5g6M}

\bibitem{mit-6s191}
MIT 6.S191.
\textit{Introduction to Deep Learning Resources}.
\url{https://introtodeeplearning.com/}

\bibitem{learnopencv-attention}
Satya Mallick.
\textit{LearnOpenCV: A Comprehensive Guide to Attention Mechanisms in Deep Learning}.
\url{https://learnopencv.com/attention-mechanism-in-neural-networks/}

\end{thebibliography}
\end{frame}