\begin{frame}[allowframebreaks]{References}
\begin{thebibliography}{99}

\bibitem{mit6s191}
MIT 6.S191: Recurrent Neural Networks (2024).\\
\url{https://introtodeeplearning.mit.edu/2024/recurrent/}

\bibitem{awsrnn}
AWS: What is an RNN?\\
\url{https://aws.amazon.com/what-is/rnn/}

\bibitem{v7labs}
V7 Labs: Complete Guide to RNNs.\\
\url{https://www.v7labs.com/blog/recurrent-neural-network}

\bibitem{mao2014deep}
Mao, J., Xu, W., Yang, Y., Wang, J., Huang, Z., \& Yuille, A. (2014). Deep Captioning with Multimodal Recurrent Neural Networks (m-RNN).\\
\textit{arXiv preprint arXiv:1412.6632}.\\
\url{https://arxiv.org/abs/1412.6632}

\bibitem{elman1990finding}
Elman, J. L. (1990). Finding structure in time.\\
\textit{Cognitive Science}, 14(2), 179-211.

\bibitem{hochreiter1997long}
Hochreiter, S., \& Schmidhuber, J. (1997). Long short-term memory.\\
\textit{Neural Computation}, 9(8), 1735-1780.

\bibitem{cho2014learning}
Cho, K., van Merriënboer, B., Gulcehre, C., Bahdanau, D., Bougares, F., Schwenk, H., \& Bengio, Y. (2014). Learning Phrase Representations using RNN Encoder–Decoder for Statistical Machine Translation.\\
\textit{arXiv preprint arXiv:1406.1078}.\\
\url{https://arxiv.org/abs/1406.1078}

\bibitem{bahdanau2014neural}
Bahdanau, D., Cho, K., \& Bengio, Y. (2014). Neural Machine Translation by Jointly Learning to Align and Translate.\\
\textit{arXiv preprint arXiv:1409.0473}.\\
\url{https://arxiv.org/abs/1409.0473}

\bibitem{xu2015show}
Xu, K., Ba, J., Kiros, R., Cho, K., Courville, A., Salakhutdinov, R., Zemel, R., \& Bengio, Y. (2015). Show, Attend and Tell: Neural Image Caption Generation with Visual Attention.\\
\textit{International Conference on Machine Learning (ICML)}.\\
\url{https://arxiv.org/abs/1502.03044}

\bibitem{vaswani2017attention}
Vaswani, A., Shazeer, N., Parmar, N., Uszkoreit, J., Jones, L., Gomez, A. N., Kaiser, Ł., \& Polosukhin, I. (2017). Attention is All You Need.\\
\textit{Advances in Neural Information Processing Systems (NeurIPS)}.\\
\url{https://arxiv.org/abs/1706.03762}

\bibitem{bengio1994learning}
Bengio, Y., Simard, P., \& Frasconi, P. (1994). Learning long-term dependencies with gradient descent is difficult.\\
\textit{IEEE Transactions on Neural Networks}, 5(2), 157-166.

\end{thebibliography}
\end{frame}