\begin{frame}[allowframebreaks]{References}

\begin{thebibliography}{9}

\bibitem{cs231n}
Fei-Fei Li, Yunzhu Li, and Ruohan Gao. 
\textit{Stanford CS231n: Deep Learning for Computer Vision}. 
\url{http://cs231n.stanford.edu/index.html}

\bibitem{dl4cv}
Assaf Shocher, Shai Bagon, Meirav Galun, and Tali Dekel. 
\textit{WAIC DL4CV Deep Learning for Computer Vision: Fundamentals and Applications}. 
\url{https://dl4cv.github.io/index.html}

\bibitem{umich}
Justin Johnson. 
\textit{UMich EECS 498.008/598.008: Deep Learning for Computer Vision}. 
\url{https://web.eecs.umich.edu/~justincj/teaching/eecs498/WI2022/}

\bibitem{dosovitskiy2020vit}
A. Dosovitskiy, L. Beyer, A. Kolesnikov, D. Weissenborn, X. Zhai, T. Unterthiner, M. Dehghani, M. Minderer, G. Heigold, S. Gelly, J. Uszkoreit, N. Houlsby,
\textit{An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale},
International Conference on Learning Representations (ICLR), 2021.

\bibitem{touvron2021deit}
H. Touvron, M. Cord, M. Douze, F. Massa, A. Sablayrolles, H. Jégou,
\textit{Training data-efficient image transformers \& distillation through attention},
International Conference on Machine Learning (ICML), 2021.

\bibitem{liu2021swin}
Z. Liu, Y. Lin, Y. Cao, H. Hu, Y. Wei, Z. Zhang, S. Lin, B. Guo,
\textit{Swin Transformer: Hierarchical Vision Transformer using Shifted Windows},
International Conference on Computer Vision (ICCV), 2021.

\bibitem{yuan2021tokens}
L. Yuan, Y. Chen, T. Wang, W. Yu, Y. Shi, Z. Jiang, F. E. Tay, J. Feng, S. Yan,
\textit{Tokens-to-Token ViT: Training Vision Transformers from Scratch on ImageNet},
International Conference on Computer Vision (ICCV), 2021.

\bibitem{carion2020detr}
N. Carion, F. Massa, G. Synnaeve, N. Usunier, A. Kirillov, S. Zagoruyko,
\textit{End-to-End Object Detection with Transformers},
European Conference on Computer Vision (ECCV), 2020.

\end{thebibliography}
\end{frame}